---
title: "ADA experiment"
author: '2150156'
date: "2023-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(dplyr)
library(summarytools)
library(rstatix)
library(car)
library(ggplot2)
library(ggpubr) 
```

## Variables' Description

- participant_number: unique ID for each participant 
- condition: one control condition and three treatment conditions
- age: age group for each participant
- education: education group for each participant
- working experience: not relevant(0), relevant(1)
- *_fin: statistics for final decision stage
- typeI_*: number of type I errors each participant 
- typeII_* number of type II errors each participant 
- ai_typeI: number of type I errors the AI made 
- ai_typeII: number of type II errors the AI made 
- confidence_*_avg : average confidence ratings each participant gave 

## Data Analysis

```{r}
## Read data
df <- read.csv("ADA_Experiment_checked.csv",header = T,stringsAsFactors = T)
```

```{r}
## Compute participant-level metrics 
df_outcomes <- df %>%
    group_by(condition, participant_number) %>%
    summarise_at(vars(
        "age","education","work_experience",
        "typeI_init", "typeII_init",
        "typeI_fin", "typeII_fin",
        "ai_typeI", "ai_typeII",
        "badloans_num", "goodloans_num",
        "confidence_init_avg", "confidence_fin_avg"
        ),
        sum, 
        na.rm = TRUE
        )

## Compute Type I (false positive) Rate  & 
## Type II (false negative) Rate 

TypeI_Rate <- function(df, typeIrate_col, typeI_col){
  df[typeIrate_col] <- df[typeI_col]/df['goodloans_num']
  return(df)
}
TypeII_Rate <- function(df, typeIIrate_col, typeII_col){
  df[typeIIrate_col] <- df[typeII_col]/df['badloans_num']
  return(df)
}

df_outcomes <- TypeI_Rate(
  df_outcomes, 'typeIrate_init', 'typeI_init')

df_outcomes <- TypeI_Rate(
  df_outcomes, 'typeIrate_fin', 'typeI_fin')

df_outcomes <- TypeI_Rate(
  df_outcomes, 'ai_typeIrate', 'ai_typeI')

df_outcomes <- TypeII_Rate(
  df_outcomes, 'typeIIrate_init', 'typeII_init')

df_outcomes <- TypeII_Rate(
  df_outcomes, 'typeIIrate_fin', 'typeII_fin')

df_outcomes <- TypeII_Rate(
  df_outcomes, 'ai_typeIIrate', 'ai_typeII')
```

```{r}
## Change data type
df_outcomes$age <- as.factor(df_outcomes$age)
df_outcomes$education <- as.factor(df_outcomes$education)
df_outcomes$work_experience <- as.factor(df_outcomes$work_experience)
```

```{r}
## Summary Statistics
perf_msrs <- c(
  "typeIrate_init", "typeIrate_fin",  
  "typeIIrate_init", "typeIIrate_fin", 
  "ai_typeIrate", "ai_typeIIrate",
  "confidence_init_avg","confidence_fin_avg")

## Summary Stats overall - all conditions
df_outcomes_descr <- dfSummary(df_outcomes[perf_msrs])

filename <- "descriptives_outcomes_pilotADA.html"

view(df_outcomes_descr, file = filename)

## Summary Stats - each condition
df_outcomes_descr_conds <- dfSummary(df_outcomes)

name <- "descriptives_outcomes_pilotADA_conds.html"

view(df_outcomes_descr_conds, file = name)
```

### First examine Type I error rate

```{r}
## look at distribution shape of my DVs for each condition
Baseline <- df_outcomes[
  df_outcomes$condition == "Baseline", ]

RA <- df_outcomes[
  df_outcomes$condition == "RA", ]

Update <- df_outcomes[
  df_outcomes$condition == "Update", ]

Feedback <- df_outcomes[
  df_outcomes$condition == "Feedback", ]

hist(Baseline$typeIrate_fin,freq = F)
lines(density(Baseline$typeIrate_fin),col="red")

hist(RA$typeIrate_fin,freq = F)
lines(density(RA$typeIrate_fin),col="red")

hist(Update$typeIrate_fin,freq = F)
lines(density(Update$typeIrate_fin),col="red")

hist(Feedback$typeIrate_fin,freq = F)
lines(density(Feedback$typeIrate_fin),col="red")
```

```{r}
## The distributions are not normal.
## Comparing between 4 conditions using Kruskal Wallis Test 
kruskal.test(
  typeIrate_fin ~ condition,
  data = df_outcomes
      )
```

```{r}
## Pairwise Comparisons
pairwise.wilcox.test( ## pairwise Mann Whitney U tests
  df_outcomes$typeIrate_fin, df_outcomes$condition,
  p.adjust.method = "BH")
```

```{r}
## Visualization(barplot)
barplot1 <- ggbarplot( 
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "typeIrate_fin", 
  ylab = "Type I Rate - Final Decisions",
  fill = "condition", ## color for each condition
  add = "mean_se", ## includes standard error bar
  label = TRUE, ## includes summary stats on chart
  label.pos = "in", 
  lab.nb.digits = 2) + ## round displayed stats to 2 decimal places
    stat_compare_means(method = "kruskal.test") + ## displays kruskal wallis results on chart 
    theme(text = element_text(size = 12, face = "bold"),
          axis.text=element_text(size = 12),
          legend.position = "none")

print(barplot1)

ggsave(plot = barplot1, width = 10, height = 8, dpi = 300,
           filename =  "barplot_pilotTypeI.jpeg")

```

```{r}
## Visualization(boxplot)
boxplot1 <- ggboxplot(
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "typeIrate_fin", ylab = "Type I Rate - Final Decisions",
  color = "condition", ## border color for each condition
  palette = c("green", "blue", "yellow", "pink"),
  order = c("Baseline", "RA", "Update", "Feedback")) + 
  stat_compare_means(method = "kruskal.test")  ## displays kruskal wallis

print(boxplot1)

ggsave(plot = boxplot1, width = 10, height = 8, dpi = 300,
           filename =  "boxplot_pilotTypeI.jpeg")
```

### Then examine Type II error rate

```{r}
## look at distribution shape of my DVs for each condition
hist(Baseline$typeIIrate_fin,freq = F)
lines(density(Baseline$typeIIrate_fin),col="red")

hist(RA$typeIIrate_fin,freq = F)
lines(density(RA$typeIIrate_fin),col="red")

hist(Update$typeIIrate_fin,freq = F)
lines(density(Update$typeIIrate_fin),col="red")

hist(Feedback$typeIIrate_fin,freq = F)
lines(density(Feedback$typeIIrate_fin),col="red")
```

```{r}
## The distributions are not normal.
## Comparing between 4 conditions using Kruskal Wallis Test 
kruskal.test(
  typeIIrate_fin ~ condition,
  data = df_outcomes
      )
```

```{r}
## Pairwise Comparisons
BR <- df_outcomes[
  df_outcomes$condition == "Baseline" | df_outcomes$condition == "RA", ]

BU <- df_outcomes[
  df_outcomes$condition == "Baseline" | df_outcomes$condition == "Update", ]

BF <- df_outcomes[
  df_outcomes$condition == "Baseline" | df_outcomes$condition == "Feedback", ]

RU <- df_outcomes[
  df_outcomes$condition == "RA" | df_outcomes$condition == "Update", ]

RF <- df_outcomes[
  df_outcomes$condition == "RA" | df_outcomes$condition == "Feedback", ]

UF <- df_outcomes[
  df_outcomes$condition == "Update" | df_outcomes$condition == "Feedback", ]

## Mann Whitney U Test (aka Wilcoxon Rank Sum test)

coin::wilcox_test(typeIIrate_fin ~ condition,
            data = BR,  
            distribution = "exact", # p = 0.02973 
            conf.int = TRUE
            )

coin::wilcox_test(typeIIrate_fin ~ condition, 
            data = BU,  
            distribution = "exact", # p = 0.1543
            conf.int = TRUE)

coin::wilcox_test(typeIIrate_fin ~ condition, 
            data = BF,   
            distribution = "exact", # p = 0.74
            conf.int = TRUE)

coin::wilcox_test(typeIIrate_fin ~ condition,
            data = RU,  
            distribution = "exact", # p = 0.7208
            conf.int = TRUE
            )

coin::wilcox_test(typeIIrate_fin ~ condition, 
            data = RF,  
            distribution = "exact", # p = 0.01785
            conf.int = TRUE)

coin::wilcox_test(typeIIrate_fin ~ condition, 
            data = UF,  
            distribution = "exact", # p = 0.08989
            conf.int = TRUE)

pvals <- c(0.02973, 0.1543, 0.74, 0.7208, 0.01785, 0.08989)
p.adjust (pvals, method="BH") # 0.08919 0.23145 0.74000 0.74000 0.08919 0.17978 
```

```{r}
## Visualization(barplot)
barplot2 <- ggbarplot( 
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "typeIIrate_fin", 
  ylab = "Type II Rate - Final Decisions",
  fill = "condition", ## color for each condition
  add = "mean_se", ## includes standard error bar
  label = TRUE, ## includes summary stats on chart
  label.pos = "in", 
  lab.nb.digits = 2) + ## round displayed stats to 2 decimal places
    stat_compare_means(method = "kruskal.test") + ## displays kruskal wallis results on chart 
    theme(text = element_text(size = 12, face = "bold"),
          axis.text=element_text(size = 12),
          legend.position = "none")

print(barplot2)

ggsave(plot = barplot2, width = 10, height = 8, dpi = 300,
           filename =  "barplot_pilotTypeII.jpeg")

```

```{r}
## Visualization(boxplot)
boxplot2 <- ggboxplot(
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "typeIIrate_fin", ylab = "Type II Rate - Final Decisions",
  color = "condition", ## border color for each condition
  palette = c("green", "blue", "yellow", "pink"),
  order = c("Baseline", "RA", "Update", "Feedback")) + 
  stat_compare_means(method = "kruskal.test")  ## displays kruskal wallis

print(boxplot2)

ggsave(plot = boxplot2, width = 10, height = 8, dpi = 300,
           filename =  "boxplot_pilotTypeII.jpeg")
```

### Finally examine average confidence

```{r}
## look at distribution shape of my DVs for each condition
hist(Baseline$confidence_fin_avg,freq = F)
lines(density(Baseline$confidence_fin_avg),col="red")

hist(RA$confidence_fin_avg,freq = F)
lines(density(RA$confidence_fin_avg),col="red")

hist(Update$confidence_fin_avg,freq = F)
lines(density(Update$confidence_fin_avg),col="red")

hist(Feedback$confidence_fin_avg,freq = F)
lines(density(Feedback$confidence_fin_avg),col="red")
```

```{r}
## The distributions are close to normal and the variances between 4 conditions are similar.
bartlett.test(list(Baseline$confidence_fin_avg, RA$confidence_fin_avg, Update$confidence_fin_avg,Feedback$confidence_fin_avg)) ## p = 0.31

## Comparing between 4 conditions using ANOVA(controlling for age, education, work_experience)
test <- lm(confidence_fin_avg ~ condition + work_experience + age + education,
  data = df_outcomes)
anova(test)
summary(test)
```


```{r}
## Visualization(barplot)
barplot3 <- ggbarplot( 
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "confidence_fin_avg", 
  ylab = "Confidence - Final Decisions",
  fill = "condition", ## color for each condition
  add = "mean_se", ## includes standard error bar
  label = TRUE, ## includes summary stats on chart
  label.pos = "in", 
  lab.nb.digits = 2) + ## round displayed stats to 2 decimal places
    stat_compare_means(method = "anova") + ## displays ANOVA results on chart 
    theme(text = element_text(size = 12, face = "bold"),
          axis.text=element_text(size = 12),
          legend.position = "none")

print(barplot3)

ggsave(plot = barplot3, width = 10, height = 8, dpi = 300,
           filename =  "barplot_pilotConfidence.jpeg")

```

```{r}
## Visualization(boxplot)
boxplot3 <- ggboxplot(
  data = df_outcomes, 
  x = "condition", xlab = "Condition",
  y = "confidence_fin_avg", ylab = "Confidence - Final Decisions",
  color = "condition", ## border color for each condition
  palette = c("green", "blue", "yellow", "pink"),
  order = c("Baseline", "RA", "Update", "Feedback")) + 
  stat_compare_means(method = "anova")  ## displays ANOVA

print(boxplot3)

ggsave(plot = boxplot3, width = 10, height = 8, dpi = 300,
           filename =  "boxplot_pilotConfidence.jpeg")
```